Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/provided_classes.tex
add filenames

\hypertarget{collection-class-routines}{%
\subsection{Collection Class Routines}\label{collection-class-routines}}

These routines are used in the definition of the collection classes
::routine CommonxXor /* Returns a new collection that contains all items
from self and the argument except that all indexes that appear in both
collections are removed. \emph{/ /} When the target is a bag, there may
be an index in the bag that is duplicated and the same value as an index
in the argument. Should one copy of the index survive in the bag?
\emph{/ vel if (arg(1)\textasciitilde class==.Set \&
arg(2)\textasciitilde class==.Bag) then v=2 if
(arg(1)\textasciitilde class==.Table \&
arg(2)\textasciitilde class==.Bag) then v=2 if
(arg(1)\textasciitilde class==.Table \&
arg(2)\textasciitilde class==.Relation) then v=2 if
(arg(1)\textasciitilde class==.Directory \&
arg(2)\textasciitilde class==.Bag) then v=2 if
(arg(1)\textasciitilde class==.Directory \&
arg(2)\textasciitilde class==.Relation) then v=2 /} This version it
does: \emph{/ if v=1 then do This = arg(1) /} self of caller */
r=This\textsubscript{class}new ab=MayEnBag (arg (2) )
ss=This\textasciitilde supplier do while ss\textasciitilde available
r{[}ss\textasciitilde index{]} =ss\textasciitilde item
ss\textasciitilde next end cs=ab\textasciitilde supplier do while
cs\textasciitilde available if r\textsubscript{hasindex(cs}index) then
r\textasciitilde remove (cs\textasciitilde index) else
r{[}cs\textasciitilde index{]} =cs\textasciitilde item
cs\textasciitilde next end return r end

/* But following matches practice on Set\textasciitilde XOR(bag) etc. */

This = arg(1) /* self of caller */

r=This\textsubscript{class}new

ab=MayEnBag (arg (2) )

ss=This\textasciitilde supplier

do while ss\textasciitilde available if
\ab\textsubscript{hasindex(ss}index) then r{[}ss\textasciitilde index{]}
=ss\textasciitilde item ss\textasciitilde next end

cs=ab\textasciitilde supplier

do while cs\textasciitilde available if
\This\textsubscript{hasindex(cs}index) then
r{[}cs\textasciitilde index{]} =cs\textasciitilde item es-next end

return r

::routine CommoniIntersect /* Returns a new collection of the same class
as SELF that contains the items from SELF that have indexes also in the
argument. \emph{/ /} Actually an index in SELF can only be `matched'
with one in the argument once. Hence copy and removal. \emph{/ This =
arg(1) /} self of caller */ w= .Bag\textasciitilde new
sc=This\textasciitilde supplier do while sc\textasciitilde available
wlsc\textasciitilde index{]} =sc\textasciitilde index sc-next end
r=This\textsubscript{class}new

cs=MayEnBag (arg(2))\textasciitilde supplier do while
cs\textasciitilde available i=cs\textasciitilde index if
w\textasciitilde hasindex(i) then do r{[}i{]}=This {[}i{]}
w\textasciitilde remove (i) end cs\textasciitilde next end return r

::routine CommonUnion /* Returns a new collection of the same class as
SELF that contains all the items from SELF and items from the argument
that have an index not in the first. \emph{/ /} Best to add them all. By
adding non-receiver first we ensure that receiver takes priority when
same indexes. \emph{/ This = arg(1) /} self of caller */
r=This\textsubscript{class}new cs=MayEnBag
(arg(2))\textasciitilde supplier do while cs\textasciitilde available
r{[}cs\textasciitilde index{]} =cs\textasciitilde item
cs\textasciitilde next end cs=This\textasciitilde supplier do while
cs\textasciitilde available r{[}cs\textasciitilde index{]}
=cs\textasciitilde item cs\textasciitilde next end return r

::routine CommonDifference /* Returns a new collection containing only
those index-item pairs from the SELF whose indexes the other collection
does not contain. \emph{/ This = arg(1) /} self of caller */
r=This\textsubscript{class}new cs=This\textasciitilde supplier do while
cs\textasciitilde available r{[}cs\textasciitilde index{]}
=cs\textasciitilde item es-next end cs=MayEnBag
(arg(2))\textasciitilde supplier do while cs\textasciitilde available
r\textasciitilde remove (cs\textasciitilde index) es-next end return r

::routine MayEnBag

/* For List and Queue the indexes are dropped. */ rearg(1) if r-clags ==
.List \textbar{} r-class == .Queue then r=EnBag(r) return r

::routine EnBag r=.Bag\textasciitilde new
s=arg(1)\textasciitilde supplier do while s\textasciitilde available if
arg(1)\textasciitilde class == .List \textbar{}
arg(1)\textasciitilde class == .Queue then r{[}s\textasciitilde item{]}
=s\textasciitilde item else /* This case is when the receiver is a Bag.
*/ r{[}s\textasciitilde index{]} =s\textasciitilde index
s\textasciitilde-next end return r

\hypertarget{the-collection-class}{%
\subsection{The collection class}\label{the-collection-class}}

::Cclass `Collection'

\hypertarget{init}{%
\subsubsection{INIT}\label{init}}

::method init

expose a /* A collection is modelled as using 3 slots in an array for
each element. The first slot holds the item, the second the index, and
the third is used by particular types of collection. This order of slots
is arbitary, chosen to match order of arguments for PUT and
SUPPLIER\textasciitilde NEW. \emph{/ /} The first set of 3 slots is
reserved for other purposes, to avoid having separate variables which
the subclassing would need to access. */

a=.array\textasciitilde new

a{[}1{]} /\emph{ItemsCount}/=0

a{[}2{]}/\emph{Unique}/=0

return self

\hypertarget{exposed}{%
\subsubsection{EXPOSED}\label{exposed}}

::method exposed private expose a

/* This method allows subclasses to get at the implementation of
Collection. */ return a

\hypertarget{findindex}{%
\subsubsection{FINDINDEX}\label{findindex}}

::method findindex private expose a /* Returns array index if the
collection contains any item associated with the index specified or
returns 0 otherwise. \emph{/ do j=4 by 3 to
1+3}a{[}1{]}/\emph{ItemsCount}/ if alj+l{]}==arg(1) then return j end j
return 0

\hypertarget{at}{%
\subsubsection{AT}\label{at}}

::method at /* vANY \emph{/ expose a /} Returns the item associated with
the specified index. */ j=self\textasciitilde findindex(arg(1)) if j=0
then return .nil return a{[}j{]}

\hypertarget{section}{%
\subsubsection{{[}{]}}\label{section}}

::method `{[}{]}' /* Synonym for the AT method. */ forward message `AT'

\hypertarget{put}{%
\subsubsection{PUT}\label{put}}

::method put /* rANY rANy \emph{/ expose a use arg item, index /}
Replaces any existing item associated with the specified index with the
new item. Otherwise adds the item-index pair. \emph{/
j=self\textasciitilde findindex (index) if j\textgreater0 then do
alj{]}=item return end a{[}1{]} /}ItemsCount\emph{/=a {[}1{]}
/}ItemsCount\emph{/+1 j=1+3}al{[}1{]} /\emph{ItemsCount}/ alj{]}=item
a{[}lj+1{]} =index a{[}j+2{]}=0 return /* Error 91 in OOI if context
requiring result. */

\hypertarget{section-1}{%
\subsubsection{{[}{]}=}\label{section-1}}

:smethod `{[}{]}=' /* Synonym for the PUT method. */ forward message
`PUT'

\hypertarget{hasindex}{%
\subsubsection{HASINDEX}\label{hasindex}}

::method hasindex /* vANY \emph{/ /} Returns 1 (true) if the collection
contains any item associated with the index specified or returns 0
(false) otherwise. */

return self\textasciitilde findindex (arg(1))\textgreater0

\hypertarget{items}{%
\subsubsection{ITEMS}\label{items}}

::method items expose a

/* Returns the number of items in the collection. \emph{/ return
a{[}1{]}/}ItemsCount*/

\hypertarget{remove}{%
\subsubsection{REMOVE}\label{remove}}

::method remove /* vANY */

expose a /* Returns and removes from a collection the member item with
the specified index. */

j=self\textasciitilde findindex(arg(1))

if j=0 then return .nil

r=al{[}j{]}

self\textasciitilde removeit (j)

return r

\hypertarget{removeit}{%
\subsubsection{REMOVEIT}\label{removeit}}

::method removeit private

expose a

use arg j

/* Remove relevant slots from the array, with compaction. */

do j=j+3 by 3 to 14+3\emph{a{[}1{]}/}ItemsCount*/
alj-3l=saljl;alj-2{]}=al{[}j+1{]};alj-ll=alj+2{]} end j

a{[}1{]} /\emph{ItemsCount}/=a {[}1{]} /\emph{ItemsCount}/-1

return

\hypertarget{makearrya}{%
\subsubsection{MAKEARRYA}\label{makearrya}}

::method makearray

expose a /* Returns a single-index array containing the receiver list
items. \emph{/ r= .array\textasciitilde new /} To build result in. */

do j=4 by 3 to 1+3\emph{a{[}1{]}/}ItemsCount*/
r{[}r\textasciitilde dimension(1)+1{]}=a{[}j{]} end j

return r

\hypertarget{makearrayx}{%
\subsubsection{MAKEARRAYX}\label{makearrayx}}

::method makearrayx private expose a /* Returns a single-index array
containing the receiver index items. \emph{/ r=
.array\textasciitilde new /} To build result in. \emph{/ do j=4 by 3 to
1+3}a{[}1{]}/\emph{ItemsCount}/
r{[}r\textasciitilde dimension(1)+1{]}=a{[}j+1{]} end j return r

\hypertarget{supplier}{%
\subsubsection{SUPPLIER}\label{supplier}}

::method supplier expose a

/* Returns a supplier object for the list. */ return
.supplier\textsubscript{new(self}makearray: .collection,
self\textasciitilde makearrayx)

\hypertarget{class-list}{%
\subsection{Class list}\label{class-list}}

::¢class `List' subclass Collection

\hypertarget{put-1}{%
\subsubsection{PUT}\label{put-1}}

::method put /* vANY rANY */ use arg item, index
a=self\textasciitilde exposed

/* PUT for a List must not be an insertion. */
j=self\textasciitilde findindex (index) if j=0 then call Raise
`Syntax',93.918 alj{]}=item return

\hypertarget{of}{%
\subsubsection{OF}\label{of}}

::method of class /* 1 or more oANY Are they omittable? Not in I00
\emph{/ /} Returns a newly created list containing the specified value
objects in the order specified. */ r= self \textasciitilde{} new do j =
1 to arg() r\textasciitilde{} insert (arg(j)) end j return r

\hypertarget{insert}{%
\subsubsection{INSERT}\label{insert}}

::method insert /* rANY oANY */

use arg item, index

a=self\textasciitilde exposed /* Returns a list-supplied index for a new
item, of specified value, which is added to the list. The new item
follows the existing item with the specified index in the list ordering.
\emph{/ /} Establish the index of what preceeds the new element. \emph{/
/} If there was no index given, the new item becomes the last on list.
\emph{/ /} .nil argument means first */

if arg(2,`E') then p=arg(2)

else p=self\textasciitilde last

/* Convert from list index to underlying array index. */

if p==.nil then j=1

else j=self\textasciitilde findindex(p)

if j=0 then call Raise `Syntax',93.918

j=j+3 /* Where new entry will be. \emph{/ /} Move space to required
place. */

a{[}1{]} /\emph{ItemsCount}/=a {[}1{]} /\emph{ItemsCount}/+1

do k=1+3\emph{a{[}1{]}/}ItemsCount*/ by -3 to j+3

a{[}k{]} =a{[}k-3{]} ;a{[}k+1{]} =a{[}k-2{]} ;a{[}k{]} =a{[}k-3{]} end

/* Insert new element. */

alj{]}=item /* A new, unique, index is needed. \emph{/ /} The basic
requirement is for something unique, so this would be correct:

is.object\textasciitilde new /* a unique object, used as a key (the
index on the list) \emph{/ }

/ /* However, a number can be used. (At risk of the user thinking it is
sensible to do arithmetic on it.) */

a{[}j+1{]}=a{[}2{]}/\emph{Unique}/;a{[}2{]} /\emph{Unique}/=a{[}2{]}
/\emph{Unique}/+1

a{[}j+2{]}=0

return a{[}j+1{]}

\hypertarget{first}{%
\subsubsection{FIRST}\label{first}}

::method first a=self\textasciitilde exposed

/* Returns the index of the first item in the list. \emph{/ if
a{[}1{]}/}ItemsCount*/=0 then return .nil

141 return a{[}5{]}

\hypertarget{last}{%
\subsubsection{LAST}\label{last}}

::smethod last a=self\textasciitilde exposed

/* Returns the index of the last item in the list. \emph{/ if
a{[}1{]}/}ItemsCount*/=0 then return .nil return
a{[}3\emph{a{[}1{]}/}ItemsCount*/+2{]}

\hypertarget{firstitem}{%
\subsubsection{FIRSTITEM}\label{firstitem}}

::method firstitem a=self\textasciitilde exposed

/* Returns the first item in the list. \emph{/ if
a{[}1{]}/}ItemsCount*/=0 then return .nil return a{[}4{]}

\hypertarget{lastitem}{%
\subsubsection{LASTITEM}\label{lastitem}}

::method lastitem a=self\textasciitilde exposed

/* Returns the last item in the list. \emph{/ if
a{[}1{]}/}ItemsCount*/=0 then return .nil return
a{[}3\emph{a{[}1{]}/}ItemsCount*/+1{]}

\hypertarget{next}{%
\subsubsection{NEXT}\label{next}}

::method next /* vANY \emph{/ a=self\textasciitilde exposed /} Returns
the index of the item that follows the list item having the specified
index. \emph{/ j=self\textasciitilde findindex(arg(1)) if j=0 then call
Raise `Syntax',93.918 j=j+3 if
j\textgreater3}a{[}1{]}/\emph{ItemsCount}/ then return .nil /* Next of
last was requested. */ return a{[}j+1{]}

\hypertarget{previous}{%
\subsubsection{PREVIOUS}\label{previous}}

::method previous /* vANY */

a=self\textasciitilde exposed /* Returns the index of the item that
precedes the list item having the specified index. */

j=self\textasciitilde findindex(arg(1))

if j=0 then call Raise `Syntax',93.918

j=j-3

if j\textless4 then return .nil /* Previous of first was requested. */

return a{[}j+1{]}

\hypertarget{section-2}{%
\subsubsection{SECTION}\label{section-2}}

::method section /* rANY oWHOLE\textgreater=0 */

a=self\textasciitilde exposed /* Returns a new list containing selected
items from the receiver list. The first item in the new list is the item
corresponding to the index specified, in the receiver list. */

j=self\textasciitilde findindex(arg(1))

if j=0 then call Raise `Syntax',93.918

r= .list\textasciitilde new /* To build result in. \emph{/ if arg(2,`E')
then s = arg(2) else s = self\textasciitilde items; do s
r\textasciitilde insert (a{[}j{]}) j=j+3 if
j\textgreater1+3}a{[}1{]}/\emph{ItemsCount}/ then leave end return r

\hypertarget{class-queue}{%
\subsection{Class queue}\label{class-queue}}

::class `Queue' subclass Collection

/* A queue is a sequenced collection with whole-number indexes. The

indexes specify the position of an item relative to the head (first
item) of the queue. Adding or removing an item changes the association
of an index to its queue item. */

\hypertarget{push}{%
\subsubsection{PUSH}\label{push}}

::method push /* vANY */

/* Adds the object value to the queue at its head. \emph{/
a=self\textasciitilde exposed a{[}1{]} /}ItemCount\emph{/=a {[}1{]}
/}ItemCount*/+1

/* Slide along to make a space. \emph{/ do
j=1+3}a{[}1{]}/\emph{ItemCount}/ by -3 to 7

aljl=alj-3{]} alj+l{]}=a{[}j-2{]}+1; /* Index changes \emph{/ end j
a{[}4{]}=arg(1) a{[}5J=1 return 12.2.4.2 PULL{]}=a{[}j+4{]}-1; /} Index
changes */ end j return r

\hypertarget{queue}{%
\subsubsection{QUEUE}\label{queue}}

::method queue /* vANY \emph{/ /} Adds the object value to the queue at
its tail. \emph{/ a=self\textasciitilde exposed a{[}1{]}
/}ItemCount\emph{/=a {[}1{]} /}ItemCount\emph{/+1 a{[}1+3\emph{a{[}1{]}
/}ItemCount*/{]}=arg(1) a{[}2+3\emph{a{[}1{]}
/}ItemCount*/{]}=a{[}1{]}/}ItemCount*/ return

\hypertarget{peek}{%
\subsubsection{PEEK}\label{peek}}

::method peek

a=self\textasciitilde exposed /* Returns the item at the head of the
queue. The collection remains unchanged. */

return a{[}4{]}

\hypertarget{remove-1}{%
\subsubsection{REMOVE}\label{remove-1}}

::method remove /* rcWHOLE\textgreater O \emph{/ /} Returns and removes
from a collection the member item with the specified index. \emph{/
a=self\textasciitilde exposed if
a{[}1{]}/}ItemCount\emph{/\textless arg(1) then return .nil
r=self\textasciitilde remove: super (arg(1)) /} Reset the indexes.
\emph{/ k=0 do j=4 by 3 to 1+3}a{[}1{]}/\emph{ItemsCount}/ k=k+1
alj+l{]}=k end j return r

\hypertarget{class-table}{%
\subsection{Class table}\label{class-table}}

::Class `Table' subclass Collection

\hypertarget{makearray}{%
\subsubsection{MAKEARRAY}\label{makearray}}

::method makearray /* Returns a single-index array containing the index
objects. \emph{/ /} This is different from Collection MAKEARRAY where
items rather than indexes are in the returned array. */ forward message
`MAKEARRAYX'

\hypertarget{union}{%
\subsubsection{UNION}\label{union}}

::method union /* rCOLLECTION */ return CommonUnion (self,arg(1))

\hypertarget{intersection}{%
\subsubsection{INTERSECTION}\label{intersection}}

::method intersection /* rCOLLECTION */ return CommoniIntersect
(self,arg(1))

\hypertarget{xor}{%
\subsubsection{XOR}\label{xor}}

::method xor /* rCOLLECTION */ return CommonXor (self,arg(1))

\hypertarget{difference}{%
\subsubsection{DIFFERENCE}\label{difference}}

::method difference /* rCOLLECTION */ return
CommonDifference(self,arg(1))

\hypertarget{subset}{%
\subsubsection{SUBSET}\label{subset}}

>>>>

\subsection{Collection Class Routines}\label{collection-class-routines}

These routines are used in the definition of the collection classes

\lstinputlisting[language=rexx,label=CommonXor.rexx,caption=CommonXor.rexx]{CommonXor.rexx}

\subsection{The collection class}\label{the-collection-class}

\lstinputlisting[language=rexx,label=collectionclass.rexx,caption=collectionclass.rexx]{collectionclass.rexx}

\subsubsection{INIT}\label{init}

\lstinputlisting[language=rexx,label=collectioninitmethod.rexx,caption=collectioninitmethod.rexx]{collectioninitmethod.rexx}

\subsubsection{EXPOSED}\label{exposed}

\lstinputlisting[language=rexx,label=collectionexposedmethod.rexx,caption=collectionexposedmethod.rexx]{collectionexposedmethod.rexx}

\subsubsection{FINDINDEX}\label{findindex}

\lstinputlisting[language=rexx,label=collectionfindindexmethod.rexx,caption=collectionfindindexmethod.rexx]{collectionfindindexmethod.rexx}

\subsubsection{AT}\label{at}

\lstinputlisting[language=rexx,label=collectionatmethod.rexx,caption=collectionatmethod.rexx]{collectionatmethod.rexx}

\subsubsection{{[}{]}}\label{section}

\lstinputlisting[language=rexx,label=collectionatalt.rexx,caption=collectionatalt.rexx]{collectionatalt.rexx}

\subsubsection{PUT}\label{put}

\lstinputlisting[language=rexx,label=collectionputmethod.rexx,caption=collectionputmethod.rexx]{collectionputmethod.rexx}

\subsubsection{{[}{]}=}\label{section-1}

\lstinputlisting[language=rexx,label=collectionputalt.rexx,caption=collectionputalt.rexx]{collectionputalt.rexx}

\subsubsection{HASINDEX}\label{hasindex}

\lstinputlisting[language=rexx,label=collectionhasindexmethod.rexx,caption=collectionhasindexmethod.rexx]{collectionhasindexmethod.rexx}

\subsubsection{ITEMS}\label{items}

\lstinputlisting[language=rexx,label=collectionsitemsmethod.rexx,caption=collectionsitemsmethod.rexx]{collectionsitemsmethod.rexx}

\subsubsection{REMOVE}\label{remove}

\lstinputlisting[language=rexx,label=collectionsremovemethod.rexx,caption=collectionsremovemethod.rexx]{collectionsremovemethod.rexx}

\subsubsection{REMOVEIT}\label{removeit}

\lstinputlisting[language=rexx,label=collectionsremoveitmethod.rexx,caption=collectionsremoveitmethod.rexx]{collectionsremoveitmethod.rexx}

\subsubsection{MAKEARRAY}\label{makearray}

\lstinputlisting[language=rexx,label=collectionsmakearraymethod.rexx,caption=collectionsmakearraymethod.rexx]{collectionsmakearraymethod.rexx}

\subsubsection{MAKEARRAYX}\label{makearrayx}

\lstinputlisting[language=rexx,label=collectionsmakearrayxmethod.rexx,caption=collectionsmakearrayxmethod.rexx]{collectionsmakearrayxmethod.rexx}

\subsubsection{SUPPLIER}\label{supplier}

\lstinputlisting[language=rexx,label=collectionssuppliermethod.rexx,caption=collectionssuppliermethod.rexx]{collectionssuppliermethod.rexx}

\subsection{Class list}\label{class-list}

\lstinputlisting[language=rexx,label=classlist.rexx,caption=classlist.rexx]{classlist.rexx}

\subsubsection{PUT}\label{put-1}

\lstinputlisting[language=rexx,label=collectionsputmethod.rexx,caption=collectionsputmethod.rexx]{collectionsputmethod.rexx}

\subsubsection{OF}\label{of}

\lstinputlisting[language=rexx,label=collectionsofmethod.rexx,caption=collectionsofmethod.rexx]{collectionsofmethod.rexx}

\subsubsection{INSERT}\label{insert}

\lstinputlisting[language=rexx,label=collectionsinsertmethod.rexx,caption=collectionsinsertmethod.rexx]{collectionsinsertmethod.rexx}

\subsubsection{FIRST}\label{first}

\lstinputlisting[language=rexx,label=collectionsfirstmethod.rexx,caption=collectionsfirstmethod.rexx]{collectionsfirstmethod.rexx}

\subsubsection{LAST}\label{last}

\lstinputlisting[language=rexx,label=collectionslastmethod.rexx,caption=collectionslastmethod.rexx]{collectionslastmethod.rexx}

\subsubsection{FIRSTITEM}\label{firstitem}

\lstinputlisting[language=rexx,label=collectionsfirstitemmethod.rexx,caption=collectionsfirstitemmethod.rexx]{collectionsfirstitemmethod.rexx}

\subsubsection{LASTITEM}\label{lastitem}

\lstinputlisting[language=rexx,label=collectionslasttitemmethod.rexx,caption=collectionslasttitemmethod.rexx]{collectionslasttitemmethod.rexx}

\subsubsection{NEXT}\label{next}

\lstinputlisting[language=rexx,label=collectionsnexttitemmethod.rexx,caption=collectionsnexttitemmethod.rexx]{collectionsnexttitemmethod.rexx}

\subsubsection{PREVIOUS}\label{previous}

\lstinputlisting[language=rexx,label=collectionsprevioustitemmethod.rexx,caption=collectionsprevioustitemmethod.rexx]{collectionsprevioustitemmethod.rexx}

\subsubsection{SECTION}\label{section-2}

\lstinputlisting[language=rexx ,label=collectionssectionmethod.rexx,caption=collectionssectionmethod.rexx]{collectionssectionmethod.rexx}

\subsection{Class queue}\label{class-queue}

\lstinputlisting[language=rexx,label=collectionsqueueclass.rexx,caption=collectionsqueueclass.rexx]{collectionsqueueclass.rexx}

\subsubsection{PUSH}\label{push}

\lstinputlisting[language=rexx,label=queuepushmethod.rexx,caption=queuepushmethod.rexx]{queuepushmethod.rexx}

\subsubsection{PULL}\label{pull}

\lstinputlisting[language=rexx,label=collectionspullmethod.rexx,caption=collectionspullmethod.rexx]{collectionspullmethod.rexx}

\subsubsection{QUEUE}\label{queue}

\lstinputlisting[language=rexx,label=queuequeuemethod.rexx,caption=queuequeuemethod.rexx]{queuequeuemethod.rexx}

\subsubsection{PEEK}\label{peek}

\lstinputlisting[language=rexx,label=queuepeekmethod.rexx,caption=queuepeekmethod.rexx]{queuepeekmethod.rexx}

\subsubsection{REMOVE}\label{remove-1}

\lstinputlisting[language=rexx,label=queueremovemethod.rexx,caption=queueremovemethod.rexx]{queueremovemethod.rexx}

\subsection{Class table}\label{class-table}

\lstinputlisting[language=rexx,label=tableclass.rexx,caption=tableclass.rexx]{tableclass.rexx}

\subsubsection{MAKEARRAY}\label{makearray-1}

\lstinputlisting[language=rexx,label=tablemakearray.rexx,caption=tablemakearray.rexx]{tablemakearray.rexx}

\subsubsection{UNION}\label{union}

\lstinputlisting[language=rexx,label=tableunionmethod.rexx,caption=tableunionmethod.rexx]{tableunionmethod.rexx}

\subsubsection{INTERSECTION}\label{intersection}

\lstinputlisting[language=rexx,label=tableintersectionmethod.rexx,caption=tableintersectionmethod.rexx]{tableintersectionmethod.rexx}

\subsubsection{XOR}\label{xor}

\lstinputlisting[language=rexx,label=tablexormethod.rexx,caption=tablexormethod.rexx]{tablexormethod.rexx}

\subsubsection{DIFFERENCE}\label{difference}

\lstinputlisting[language=rexx,label=tabledifferencemethod.rexx,caption=tabledifferencemethod.rexx]{tabledifferencemethod.rexx}

\subsubsection{SUBSET}\label{subset}

\lstinputlisting[language=rexx,label=tablesubsetmethod.rexx,caption=tablesubsetmethod.rexx]{tablesubsetmethod.rexx}

\subsubsection{Class set}\label{class-set}

\lstinputlisting[language=rexx,label=classset.rexx,caption=classset.rexx]{classset.rexx}

\subsubsection{PUT}\label{put-2}

\lstinputlisting[language=rexx,label=setputmethod.rexx,caption=setputmethod.rexx]{setputmethod.rexx}

\subsubsection{OF}\label{of-1}

\lstinputlisting[language=rexx,label=setofmethod.rexx,caption=setofmethod.rexx]{setofmethod.rexx}

\subsubsection{UNION}\label{union-1}

\lstinputlisting[language=rexx,label=setunion.rexx,caption=setunion.rexx]{setunion.rexx}

\subsubsection{INTERSECTION}\label{intersection-1}

\lstinputlisting[language=rexx,label=setintersectionmethod.rexx,caption=setintersectionmethod.rexx]{setintersectionmethod.rexx}

\subsubsection{XOR}\label{xor-1}

\lstinputlisting[language=rexx,label=setxormethod.rexx,caption=setxormethod.rexx]{setxormethod.rexx}

\subsubsection{DIFFERENCE}\label{difference-1}

\lstinputlisting[language=rexx,label=setdifferencemethod.rexx,caption=setdifferencemethod.rexx]{setdifferencemethod.rexx}

\subsection{Class relation}\label{class-relation}

\lstinputlisting[language=rexx,label=relationclass.rexx,caption=relationclass.rexx]{relationclass.rexx}

\subsubsection{PUT}\label{put-3}

\lstinputlisting[language=rexx,label=relputmethod.rexx,caption=relputmethod.rexx]{relputmethod.rexx}

\subsubsection{ITEMS}\label{items-1}

\lstinputlisting[language=rexx,label=relitemsmethod.rexx,caption=relitemsmethod.rexx]{relitemsmethod.rexx}

\subsubsection{MAKEARRAY}\label{makearray-2}

\lstinputlisting[language=rexx,label=relmakearray method,caption=relmakearray method]{relmakearray method}

\subsubsection{SUPPLIER}\label{supplier-1}

\lstinputlisting[language=rexx,label=relsuppliermethod.rexx,caption=relsuppliermethod.rexx]{relsuppliermethod.rexx}

\subsubsection{UNION}\label{union-2}

\lstinputlisting[language=rexx,label=relunionmethod.rexx,caption=relunionmethod.rexx]{relunionmethod.rexx}

\subsubsection{INTERSECTION}\label{intersection-2}

\lstinputlisting[language=rexx,label=relintersectionmethod.rexx,caption=relintersectionmethod.rexx]{relintersectionmethod.rexx}

\subsubsection{XOR}\label{xor-2}

\lstinputlisting[language=rexx,label=relxormethod.rexx,caption=relxormethod.rexx]{relxormethod.rexx}

\subsubsection{DIFFERENCE}\label{difference-2}

\lstinputlisting[language=rexx,label=reldifferencemethod.rexx,caption=reldifferencemethod.rexx]{reldifferencemethod.rexx}

\subsubsection{SUBSET}\label{subset-1}

\lstinputlisting[language=rexx,label=relsubsetmethod.rexx,caption=relsubsetmethod.rexx]{relsubsetmethod.rexx}

\subsubsection{REMOVEITEM}\label{removeitem}

\lstinputlisting[language=rexx,label=relremoveitemmethod.rexx,caption=relremoveitemmethod.rexx]{relremoveitemmethod.rexx}

\subsubsection{INDEX}\label{index}

\lstinputlisting[language=rexx,label=relindexmethod.rexx,caption=relindexmethod.rexx]{relindexmethod.rexx}

\subsubsection{ALLAT}\label{allat}

\lstinputlisting[language=rexx,label=relAllAtmethod.rexx,caption=relAllAtmethod.rexx]{relAllAtmethod.rexx}

\subsubsection{HASITEM}\label{hasitem}

\lstinputlisting[language=rexx,label=relhasitemmethod.rexx,caption=relhasitemmethod.rexx]{relhasitemmethod.rexx}

\subsubsection{ALLINDEX}\label{allindex}

\lstinputlisting[language=rexx,label=relallindexmethod.rexx,caption=relallindexmethod.rexx]{relallindexmethod.rexx}

\lstinputlisting[language=rexx,label=bagclass.rexx,caption=bagclass.rexx]{bagclass.rexx}

\subsubsection{OF}\label{of-2}

\lstinputlisting[language=rexx,label=bagofmethod.rexx,caption=bagofmethod.rexx]{bagofmethod.rexx}

\subsubsection{PUT}\label{put-4}

\lstinputlisting[language=rexx,label=bagputmethod.rexx,caption=bagputmethod.rexx]{bagputmethod.rexx}

\subsubsection{UNION}\label{union-3}

\lstinputlisting[language=rexx,label=bagunionmethod.rexx,caption=bagunionmethod.rexx]{bagunionmethod.rexx}

\subsubsection{INTERSECTION}\label{intersection-3}

\lstinputlisting[language=rexx,label=bagintersectionmethod.rexx,caption=bagintersectionmethod.rexx]{bagintersectionmethod.rexx}

\subsubsection{XOR}\label{xor-3}

\lstinputlisting[language=rexx,label=bagxormethod.rexx,caption=bagxormethod.rexx]{bagxormethod.rexx}

\subsubsection{DIFFERENCE}\label{difference-3}

\lstinputlisting[language=rexx,label=bagdifferencemethod.rexx,caption=bagdifferencemethod.rexx]{bagdifferencemethod.rexx}

\subsection{The directory class}\label{the-directory-class}

\lstinputlisting[language=rexx,label=directoryclass.rexx,caption=directoryclass.rexx]{directoryclass.rexx}

\subsubsection{AT}\label{at-1}

\lstinputlisting[language=rexx,label=diratmethod.rexx,caption=diratmethod.rexx]{diratmethod.rexx}

\subsubsection{PUT}\label{put-5}

\lstinputlisting[language=rexx,label=dirputmethod.rexx,caption=dirputmethod.rexx]{dirputmethod.rexx}

\subsubsection{MAKEARRAY}\label{makearray-3}

\lstinputlisting[language=rexx,label=diremakearray.rexx,caption=diremakearray.rexx]{diremakearray.rexx}

\subsubsection{SUPPLIER}\label{supplier-2}

\lstinputlisting[language=rexx,label=dirsuppliermethod.rexx,caption=dirsuppliermethod.rexx]{dirsuppliermethod.rexx}

\subsubsection{UNION}\label{union-4}

\lstinputlisting[language=rexx,label=dirunionmethod.rexx,caption=dirunionmethod.rexx]{dirunionmethod.rexx}

\subsubsection{INTERSECTION}\label{intersection-4}

\lstinputlisting[language=rexx,label=dirintersectionmethod.rexx,caption=dirintersectionmethod.rexx]{dirintersectionmethod.rexx}

\subsubsection{XOR}\label{xor-4}

\lstinputlisting[language=rexx,label=dirxormethod.rexx,caption=dirxormethod.rexx]{dirxormethod.rexx}

\subsubsection{DIFFERENCE}\label{difference-4}

\lstinputlisting[language=rexx,label=dirdifferencemethod.rexx,caption=dirdifferencemethod.rexx]{dirdifferencemethod.rexx}

\subsubsection{SUBSET}\label{subset-2}

\lstinputlisting[language=rexx,label=dirsubsetmethod.rexx,caption=dirsubsetmethod.rexx]{dirsubsetmethod.rexx}

\subsubsection{SETENTRY}\label{setentry}

\lstinputlisting[language=rexx,label=dirsetentrymethod.rexx,caption=dirsetentrymethod.rexx]{dirsetentrymethod.rexx}

\subsubsection{ENTRY}\label{entry}

\lstinputlisting[language=rexx,label=direntrymethod.rexx,caption=direntrymethod.rexx]{direntrymethod.rexx}

\subsubsection{HASENTRY}\label{hasentry}

\lstinputlisting[language=rexx,label=dirhasentrymethod.rexx,caption=dirhasentrymethod.rexx]{dirhasentrymethod.rexx}

\subsubsection{SETMETHOD}\label{setmethod}

\lstinputlisting[language=rexx,label=dirsetmethod.rexxx,caption=dirsetmethod.rexxx]{dirsetmethod.rexxx}

\subsubsection{UNKNOWN}\label{unknown}

\lstinputlisting[language=rexx,label=dirunknownmethod.rexx,caption=dirunknownmethod.rexx]{dirunknownmethod.rexx}

\subsection{The stem class}\label{the-stem-class}

\emph{For some reason, the stem class doesn't have PUT and AT methods,
which stops us having a general rule about {[}{]} synonyms AT, {[}{]}=
synonyms PUT.}

\emph{Anyway, committee doing without this class as such.}

\emph{Here is temporary stuff showing how to use algebra in the
collection coding.}

\lstinputlisting[language=rexx,label=stemclass.rexx,caption=stemclass.rexx]{stemclass.rexx}

{[}The following code looks like a partial duplicate of the previous.
Formatting pending -- JMB{]}

\lstinputlisting[language=rexx,label=setlikeoperations.rexx,caption=setlikeoperations.rexx]{setlikeoperations.rexx}

\section{The stream class}\label{the-stream-class}

The stream class provides input/output on external streams.

\lstinputlisting[language=rexx,label=streamclass.rexx,caption=streamclass.rexx]{streamclass.rexx}

Initializes a stream object for a stream named name, but does not open
the stream.

\lstinputlisting[language=rexx,label=streamquerymethod.rexx,caption=streamquerymethod.rexx]{streamquerymethod.rexx}

\emph{There is also QUERY as command with method COMMAND.}

Used with options, the QUERY method returns specific information about a
stream.

\lstinputlisting[language=rexx,label=streamiomethods.rexx,caption=streamiomethods.rexx]{streamiomethods.rexx}

Returns a string after performing the specified stream command.

\lstinputlisting[language=rexx,label=steamopenmethod.rexx,caption=steamopenmethod.rexx]{steamopenmethod.rexx}

\emph{There is also OPEN as command with method COMMAND.}
